---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Loading the data:

```{r, echo = TRUE}
activity <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

* Total number of steps taken per day

```{r, echo = TRUE}
steps_per_day<-sapply(split(activity$steps, activity$date), sum)
steps_per_day
```

* Histogram of the total number of steps taken each day

```{r, echo = TRUE}
steps_per_day<-steps_per_day[complete.cases(steps_per_day)]
hist(steps_per_day, ylim=c(0,35), xlab =  "Steps per days", main = "Steps per day distribution")
text(mean(steps_per_day)-4000,33,paste("Mean = ",round(mean(steps_per_day),0)," -->"), col="blue" )
text(median(steps_per_day)+4300,33,paste("<-- Median = ",round(median(steps_per_day))), col="red" )
abline(v=mean(steps_per_day), col="blue")
abline(v=median(sort(steps_per_day)), col="red")
```

* Mean and median of the total number of steps taken per day

```{r, echo = TRUE}
mean(steps_per_day)
median(sort(steps_per_day))
```

As you can see, the mean and median are practically equal.

## What is the average daily activity pattern?

* Time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r, echo = TRUE}
steps_interval_means<-sapply(split(activity$steps, activity$interval), mean, na.rm = T)
plot(unique(activity$interval), steps_interval_means, 
     xlab="Intervals", ylab="Step", 
     main="Averaged Steps per Intervals", type="l")
```

* Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

The maximun number of steps on average is

```{r, echo = TRUE}
max(steps_interval_means)
```

that belong to the interval:

```{r, echo = TRUE}
 names(steps_interval_means[steps_interval_means==max(steps_interval_means)])
```

## Imputing missing values

* The total number of missing values in the dataset is:

```{r, echo = TRUE}
sum(!complete.cases(activity))
```

* Filling missing values with the mean of corresponding interval, in a new dataset:

```{r, echo = TRUE}
activity1<-activity
missing_values <- !complete.cases(activity1)
activity_na <-activity1[missing_values,]
for(i in activity_na$interval){  
 activity_na[activity_na$interval==i,]$steps=steps_interval_means[as.character(i)]  
}
#New dataset activity1 without missing values
activity1[missing_values,]$steps = round(activity_na$steps,2)
```

* Total number of steps taken per day:

```{r, echo = TRUE}
steps_per_day1<-sapply(split(activity1$steps, activity1$date), sum)
```

* Histogram of the total number of steps taken each day

```{r, echo = TRUE}
hist(steps_per_day1, ylim=c(0,40), xlab =  "Steps per days", main = "Steps per day distribution")
text(mean(steps_per_day1)-4000,38,paste("Mean = ",round(mean(steps_per_day1),0)," -->"), col="blue" )
text(median(steps_per_day1)+4300,38,paste("<-- Median = ",round(median(steps_per_day1))), col="red" )
abline(v=mean(steps_per_day1), col="blue")
abline(v=median(sort(steps_per_day1)), col="red")
```

* Mean and median of the total number of steps taken per day with filled missing values:

```{r, echo = TRUE}
mean(steps_per_day1)
median(sort(steps_per_day1))
```

Although we replaced the missing values with the mean of the steps per interval and as it can be seen in the plot, where the frecuency only had a little change at the portion 10000-15000, the differences between mean and median with and without missing values is very small.

So, we can conclude that there are not significant differences between both dataset.

## Are there differences in activity patterns between weekdays and weekends?

```{r, echo = TRUE}
wd<-c("Monday"="weekday","Tuesday"="weekday","Wednesday"="weekday",
      "Thursday"="weekday","Friday"="weekday", "Saturday"="weekend" ,"Sunday"="weekend")
days <- weekdays(  as.Date(activity1$date))
activity1$day_type=factor(wd[days])

library(ggplot2)
steps_means <- aggregate(activity1$steps, 
                         by=list(interval=activity1$interval, 
                                 day_type=activity1$day_type),
                         mean)
qplot( interval, x, xlab="Intervals", ylab="Steps", 
       data=steps_means,colours=c("blue", "red"),  
       geom="line")+ facet_wrap(~day_type, ncol = 1)
```

In above plot we can see significant averages changes in the patterns.

In weekends the quantity of steps, between intervals 500 and 1000, experiment a great reduction, from approximately 76 to 45 steps per interval,   
```{r, echo = TRUE}
#steps for 500 to 1000 intervals
#weekdays average steps
mean(steps_means[steps_means$day_type =="weekday" 
                 & steps_means$interval>=500 
                 & steps_means$interval<=1000, ]$x)
#weekends average steps
mean(steps_means[steps_means$day_type =="weekend" 
                 & steps_means$interval>=500 
                 & steps_means$interval<=1000, ]$x)
```

and an increment between intervals 1000 and 2000 from approximately 43 to 70 steps per interval.

```{r, echo = TRUE}
#steps for 1000 to 2000 intervals
#weekdays average steps
mean(steps_means[steps_means$day_type =="weekday" 
                 & steps_means$interval>=1000 
                 & steps_means$interval<=2000, ]$x)
#weekends average steps
mean(steps_means[steps_means$day_type =="weekend" 
                 & steps_means$interval>=1000 
                 & steps_means$interval<=2000, ]$x)
```



